Improved Approximation Algorithms for the Maximum Carpool Matching Problme

### 2016-12-19 src/abstract.tex
intro

The \textsc{\CARPOOL{}} problem is a star packing problem in directed graphs.
Formally, given a directed graph $G = (V, A)$,
a capacity function $c: V \rightarrow \N$,
and a profit function $p : A \rightarrow \R $,
a feasible \emph{carpool matching} is a subset of arcs, 
$M \subseteq A$, where
$\forall v \in V, 
\din[M](v) \cdot \dout[M](v) = 0, 
\din[M] \leq c(v),
\dout[M] \leq 1$. 
In the \textsc{\CARPOOL{}} problem we seek for a matching $M$
that maximizes the total profit of $M$.

@@

we show that the problem can be formulated as
unconstarints submodular maximization problem, 
thus it has a $\frac{1}{2}$-approximation algorithm.
For the unweighted variant of the problem when the maximum possible capacity, $C$, is
fixed, we give a $(\frac{C + 1}{2C} - \varepsilon)$-approximation algorithm
for the any $\varepsilon > 0$.

We also consider a variant of the \textsc{\CARPOOL{}} problem where every arc
has also an integral weight $w:A \to \N$ and the constraint $\din[M] \leq c(v)$
is replaced with the constraint $\sum_{(u,v)}w(u,v) \leq c(v)$.
We show that this variant of the problem can not be formulated as an
unconstarints submodular maximization as before.
We give, however, a ($2 + \varepsilon$)-approximation algorithm for this
variant, and argue that the local search algorithm can be generalized to this
variant of the problem when $C$ is fixed.

>>>>

The \textsc{\CARPOOL{}} problem is a star packing problem in directed graphs.
Formally, given a directed graph $G = (V, A)$,
a capacity function $c: V \rightarrow \N$,
and a weight function $w : A \rightarrow \R $,
a feasible \emph{carpool matching} is a subset of arcs, 
$M \subseteq A$, where
$\forall v \in V, 
\din[M](v) \cdot \dout[M](v) = 0, 
\din[M] \leq c(v),
\dout[M] \leq 1$. 
In the \textsc{\CARPOOL{}} problem we seek for a matching $M$
that maximizes the total weight of $M$.

@@

we show that the problem can be formulated as
unconstrained submodular maximization problem, 
thus it has a $\frac{1}{2}$-approximation algorithm.
For the unweighted variant of the problem when the maximum possible capacity, $C$, is
fixed, we give a $(\frac{C + 1}{2C} - \varepsilon)$-approximation algorithm
for the any $\varepsilon > 0$.

We also consider a variant of the \textsc{\CARPOOL{}} problem where every arc
has also an integral size $s:A \to \N$ and the constraint $\din[M] \leq c(v)$
is replaced with the constraint $\sum_{(u,v)}s(u,v) \leq c(v)$.
We show that this variant of the problem can not be formulated as an
unconstrained submodular maximization as before.
We give, however, a ($2 + \varepsilon$)-approximation algorithm for this
variant, and argue that the local search algorithm can be generalized to this
variant of the problem when $C$ is fixed.

### 2016-12-19 src/intro.tex
intro

A feasible \emph{carpool matching} (matching) is a triple 
$(P, D, M)$, where $P$ and $D$ form a partition of $V$, 
and $M$ is a subset of $A \cap (P \times D)$,
under the constraints that for every driver $d \in D$, 
$\din(d) \leq c(d)$, 
and for every passenger $p \in P$, ${\dout(p) \leq 1}$.
In the \textsc{\CARPOOL{}} problem we seek for a matching $(P, D, M)$ that maximizes the
total weight of $M$.
In other words, the \textsc{\CARPOOL{}} problem is about finding a set of 
(directed toward the center) vertex disjoint stars 
that maximizes the total weights on the arcs.
Figure~\ref{fig:carpool} is an example of the \textsc{\CARPOOL{}} problem.

@@

In section~\ref{sec:sub} we show that the problem can be formulated as
unconstarints submodular maximization problem, thus it has a 2-approximation
algorithm.
In section~\ref{sec:local} we consider the unweighted variant of the problem.
We show that when the maximum possible capacity, $C$, is fixed, there is a 
$(\frac{C + 1}{2C} - \varepsilon)$-approximation algorithm for the any $\varepsilon > 0$.

>>>>

a feasible \emph{carpool matching} is a subset of arcs, 
$M \subseteq A$, where
$\forall v \in V, 
\din[M](v) \cdot \dout[M](v) = 0, 
\din[M] \leq c(v),
\dout[M] \leq 1$. 
In the \textsc{\CARPOOL{}} problem we seek for a matching $M$
that maximizes the total profit of $M$.
In other words, the \textsc{\CARPOOL{}} problem is about finding a set of 
(directed toward the center) vertex disjoint stars 
that maximizes the total weights on the arcs.
Figure~\ref{fig:carpool} is an example of the \textsc{\CARPOOL{}} problem.

@@

In Section~\ref{sec:sub} we show that the problem can be formulated as
unconstrained submodular maximization problem, thus it has a
$\frac{1}{2}$-approximation algorithm.
In Section~\ref{sec:local} we consider the unweighted variant of the problem.
We show that when the maximum possible capacity, $C$, is fixed, there is a 
$(\frac{C + 1}{2C} - \varepsilon)$-approximation algorithm for the any $\varepsilon > 0$.
We also show that the problem remains APX-hard even if $C \leq 3$.
In Section~\ref{sec:group} we consider a variant of the problem, the
\textsc{group \CARPOOL} problem, where each edge has an integral size and 
the constraint $\din[M] \leq c(v)$ is replaced with the constraint
$\sum_{(u,v)}s(u,v) \leq c(v)$.
We show that this variant of the problem can not be formulated as unconstrained
submodular maximization problem as before.
We show, however, that this variant of the problem still admits a
$(\frac{1}{2} - \varepsilon)$-approximation algorithm, and argue that the local
search algorithm generalized to this variant as well.
